### 2018-08-12 documentation/paper.tex
Initial Commit

\documentclass{article}
\usepackage[style=ieee]{biblatex}
\bibliography{references.bib}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\title{Lennard-Jones Energy Minimisation}
\author{Bennett Gale}
\date{July 2018}
\lstset{language=C++, basicstyle=\ttfamily}

\begin{document}
\maketitle
\tableofcontents
\pagebreak

\section{Introduction}

The model molecular energy minimisation problem involves the geometric
configuration of a cluster of atoms in two dimensions, each connected by rigid
bonds of unit length, such that the total potential energy yielded by the sum of
individual Lennard-Jones potentials is globally minimised. The total potential
energy of the cluster is given by 
$$V(r)=\sum^{N-1}_{i=1}\sum^{N}_{j=i+1}\left(r^{-12}_{ij}-2r^{-6}_{ij}\right)$$
Where $r$ is the euclidean distance between a pair of atoms and $N$ is the
number of atoms in the chain.

The energy minimisation problem is historically difficult because of the
exponential relationship between the locally stable structures and the number of
atoms, $N$, in the molecule. That is, as the size of the cluster increases, the
number of local minima increases exponentially such that when $N$ approaches
100, there exist more than $10^{140}$ minima \cite{DAVEN1996195}. It is
therefore necessary to employ sophisticated search algorithms in order to make
larger input sizes attainable.

Section~\ref{litreview} will discuss some of the research that has already taken
place in this area. Section~\ref{proposed} will detail the algorithm proposed by
this paper.

\section{Literature Review} \label{litreview}

Biased modelling approaches, in which prior assumptions are made regarding the
structure of optimal configurations, have been used successfully in the past to
yield powerful results. Initially, for example, global minima for many of
$N\leq150$ were found using lattices derived from icosahedron
\cite{doi:10.1063/1.453492}. However while these methods can be the most
powerful, they suffer in cases when the assumed model is largely different from
the real configuration \cite{doi:10.1002/jcc.20096}. Therefore, more general,
less biased algorithms have been applied in order to increase robustness against
such exceptional cases.

Genetic algorithms have been employed in in order to successfully extend the
range of input size $N$ that is reachable through structural optimisation
\cite{DAVEN1996195, doi:10.1002/qua.560440214,PULLAN1998331}. This population
based approach emulates the Darwinian process of natural selection, and works by
performing crossover operations on pairs of candidate configurations over many
generations until a global minimum is produced. This method has been successful
in producing results that were previously unseen in literature for certain $N$
sizes, due to its generality when compared with more biased approaches.

The Conformational Space Annealing Method is a more recent unbiased global
optimisation algorithm that has produced very good results
\cite{PhysRevLett.91.080201}. This technique is highly robust against random
initial configurations thanks to its generality. As a result, it has therefore
been able to produce all global minima for $n\leq183$ without suffering from
exceptional cases like those experienced by the biased modelling algorithms.
This algorithm combines concepts of Monte Carlo with Minimisation, Genetic
Algorithm and simulated Annealing, and requires no specific knowledge or
assumptions regarding the optimal structure. Local minimisation is performed on
a set of configurations, and then similarly to the mating process of genetic
algorithm, a subset of configurations is produced. A parameter $D_cut$ is
introduced and acts as a temperature variable in the annealing process. This
parameter represents a distance measure between configurations and is used to
maintain diversity in the population \cite{PhysRevLett.91.080201}.

Another unbiased global optimisation algorithm, known as dynamic lattice
searching (DLS), works by repeatedly constructing and searching a dynamic
lattice until a global minimum is produced \cite{doi:10.1002/jcc.20096}. This
lattice is built by searching the possible location sites for added atoms, which
are then iteratively relocated from lattice sites of higher energy to those of
lower energy. This method is very efficient because of how much it reduces the
search space, but remains unbiased and therefore does not exhibit the same
limitations seen in the biased modelling based techniques.

\section{Proposed Optimisation Algorithm} \label{proposed}

\pagebreak

\printbibliography

\end{document}

### 2018-08-12 main.cpp
Added plot script

@@ -1,6 +1,43 @@
 #include <iostream>
+#include <vector>
+#include <random>
+#include <cmath>
 
-int main() {
-	std::cout << "Hello, World!" << std::endl;
+struct Molecule {
+	unsigned length;
+	std::vector<double> bond_angles;
+
+	explicit Molecule(unsigned length)
+	{
+		this->length = length;
+		bond_angles.resize(length);
+	}
+};
+
+/**
+ * This function is completely random and does not consider knots
+ */
+Molecule rand_mol(unsigned len)
+{
+	Molecule s(len);
+	std::random_device rd;
+	std::mt19937 gen(rd()); // New Mersenne-Twister generator seeded with rd
+	std::uniform_real_distribution<double> dis(-M_PI, M_PI);
+	for (auto &i : s.bond_angles)
+		i = dis(gen);
+	return s;
+}
+
+void print_mol(Molecule &s)
+{
+	for (auto &i : s.bond_angles)
+		std::cout << i << ' ';
+	std::cout << std::endl;
+}
+
+int main()
+{
+	Molecule m = rand_mol(30);
+	print_mol(m);
 	return 0;
 }

### 2018-08-12 plot.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python3
+
+from sys import stdin
+import matplotlib.pyplot as plt
+from math import cos, sin
+
+
+def plot(angle_chain, fig_name):
+    plt.figure()
+    plt.axis('equal')
+    chain = [float(x) for x in angle_chain.split()]
+    x = [0]
+    y = [0]
+    cumulative = 0
+    for alpha in chain:
+        cumulative += alpha
+        x.append(x[-1] + cos(cumulative))
+        y.append(y[-1] + sin(cumulative))
+    plt.plot(x, y)
+    plt.savefig("./documentation/figure " + fig_name + ".png")
+    print(chain)
+    plt.close()
+
+
+if __name__ == "__main__":
+    count = 0
+    for line in stdin:
+        plot(line, str(count))
+        count += 1

### 2018-08-12 main.cpp
Switched to CSV format for printing

@@ -28,16 +28,19 @@ Molecule rand_mol(unsigned len)
 	return s;
 }
 
-void print_mol(Molecule &s)
+void print_csv(Molecule& s)
 {
-	for (auto &i : s.bond_angles)
-		std::cout << i << ' ';
+	for (auto i = 0; i < s.bond_angles.size(); i++) {
+		std::cout << s.bond_angles[i];
+		if (i != s.bond_angles.size() - 1)
+			std::cout << ",";
+	}
 	std::cout << std::endl;
 }
 
 int main()
 {
-	Molecule m = rand_mol(30);
-	print_mol(m);
+	Molecule m = rand_mol(8);
+	print_csv(m);
 	return 0;
 }

### 2018-08-12 plot.py

@@ -8,7 +8,7 @@ from math import cos, sin
 def plot(angle_chain, fig_name):
     plt.figure()
     plt.axis('equal')
-    chain = [float(x) for x in angle_chain.split()]
+    chain = [float(x) for x in angle_chain.split(',')]
     x = [0]
     y = [0]
     cumulative = 0
@@ -16,14 +16,14 @@ def plot(angle_chain, fig_name):
         cumulative += alpha
         x.append(x[-1] + cos(cumulative))
         y.append(y[-1] + sin(cumulative))
-    plt.plot(x, y)
+    plt.plot(x, y, 'o-')
     plt.savefig("./documentation/figure " + fig_name + ".png")
-    print(chain)
     plt.close()
 
 
 if __name__ == "__main__":
     count = 0
     for line in stdin:
+        print(line)
         plot(line, str(count))
         count += 1

### 2018-08-13 main.cpp
Implemented Lennard-Jones potential function.

@@ -10,8 +10,12 @@ struct Molecule {
 	explicit Molecule(unsigned length)
 	{
 		this->length = length;
-		bond_angles.resize(length);
+		bond_angles.resize(length - 1);
 	}
+};
+
+struct Cartesian {
+    double x, y;
 };
 
 /**
@@ -22,9 +26,9 @@ Molecule rand_mol(unsigned len)
 	Molecule s(len);
 	std::random_device rd;
 	std::mt19937 gen(rd()); // New Mersenne-Twister generator seeded with rd
-	std::uniform_real_distribution<double> dis(-M_PI, M_PI);
-	for (auto &i : s.bond_angles)
-		i = dis(gen);
+	std::uniform_real_distribution<double> uniform(-M_PI, M_PI);
+	for (auto& i : s.bond_angles)
+		i = uniform(gen);
 	return s;
 }
 
@@ -38,9 +42,37 @@ void print_csv(Molecule& s)
 	std::cout << std::endl;
 }
 
+/**
+ * Calculate the Lennard-Jones potential energy of the given system s
+ */
+double lj_potential(Molecule& s)
+{
+    std::vector<Cartesian> cart(s.length);
+    cart[0] = {0, 0};
+    double phi = 0;
+    for (int i = 0; i < s.bond_angles.size(); i++) {
+        phi += s.bond_angles[i];
+        double x = cart[i].x + cos(phi);
+        double y = cart[i].y + sin(phi);
+	    cart[i + 1] = {x, y};
+    }
+    double v = 0;
+    for (int i = 0; i < s.length - 1; i++) {
+        for (int j = i + 1; j < s.length; j++) {
+        	double delta_x = cart[j].x - cart[i].x;
+        	double delta_y = cart[j].y - cart[i].y;
+            v += 1 / pow(pow(delta_x, 2) + pow(delta_y, 2), 6) - 2 / pow(pow(delta_x, 2) + pow(delta_y, 2), 3);
+        }
+    }
+    return v;
+}
+
 int main()
 {
-	Molecule m = rand_mol(8);
+	Molecule m(4);
+	std::vector<double> angles = {0, M_PI / 4.0, -M_PI / 2.0};
+	m.bond_angles = angles;
 	print_csv(m);
+	std::cout << lj_potential(m) << std::endl;
 	return 0;
 }

### 2018-08-13 main.cpp
Fixed error

@@ -10,7 +10,7 @@ struct Molecule {
 	explicit Molecule(unsigned length)
 	{
 		this->length = length;
-		bond_angles.resize(length - 1);
+		bond_angles.resize(length - 2);
 	}
 };
 
@@ -61,7 +61,8 @@ double lj_potential(Molecule& s)
         for (int j = i + 1; j < s.length; j++) {
         	double delta_x = cart[j].x - cart[i].x;
         	double delta_y = cart[j].y - cart[i].y;
-            v += 1 / pow(pow(delta_x, 2) + pow(delta_y, 2), 6) - 2 / pow(pow(delta_x, 2) + pow(delta_y, 2), 3);
+            v += 1 / pow(pow(delta_x, 2) + pow(delta_y, 2), 6)
+                 - 2 / pow(pow(delta_x, 2) + pow(delta_y, 2), 3);
         }
     }
     return v;

### 2018-08-13 main.cpp
Added liblbfgs library for local optimisation

@@ -2,6 +2,7 @@
 #include <vector>
 #include <random>
 #include <cmath>
+#include <lbfgs.h>
 
 struct Molecule {
 	unsigned length;

### 2018-08-15 documentation/paper.tex
Completed literature review version 2

@@
 
-Biased modelling approaches, in which prior assumptions are made regarding the
-structure of optimal configurations, have been used successfully in the past to
-yield powerful results. Initially, for example, global minima for many of
-$N\leq150$ were found using lattices derived from icosahedron
-\cite{doi:10.1063/1.453492}. However while these methods can be the most
-powerful, they suffer in cases when the assumed model is largely different from
-the real configuration \cite{doi:10.1002/jcc.20096}. Therefore, more general,
-less biased algorithms have been applied in order to increase robustness against
-such exceptional cases.
+Genetic algorithms have been applied \cite{doi:10.1002/qua.560440214,
+PULLAN1998331} successfully in order to extend the range of input size $N$ that
+is reachable through structural optimisation. This population based approach
+emulates the Darwinian process of natural selection, and works by performing
+crossover operations on pairs of candidate configurations with selection
+probabilities based on their fitness, producing new child nodes which contain
+genetic material from both parents. With every crossover operation there is a
+small chance of subsequent random mutation of the child configurations, the
+purpose of which is to promote genetic diversity within the population and
+prevent premature convergence. Over many generations the population converges
+towards a global minimum as each configuration communicates
+information about the global search space through the crossover operation. This
+method has been successful in producing results that were previously unseen in
+literature for certain input sizes.
 
-Genetic algorithms have been employed in in order to successfully extend the
-range of input size $N$ that is reachable through structural optimisation
-\cite{DAVEN1996195, doi:10.1002/qua.560440214,PULLAN1998331}. This population
-based approach emulates the Darwinian process of natural selection, and works by
-performing crossover operations on pairs of candidate configurations over many
-generations until a global minimum is produced. This method has been successful
-in producing results that were previously unseen in literature for certain $N$
-sizes, due to its generality when compared with more biased approaches.
+Judson \textit{et al.} compare the performance of GAs with simulated annealing,
+a stochastic global optimisation technique in which Monte Carlo steps are taken
+and new states are accepted if they are fitter, or with a probability based on a
+function of a temperature value \cite{doi:10.1002/qua.560440214}. When applied
+to the molecular conformation problem, they found that the GA method and the SA
+method, both applied as global meta-heuristics in conjunction with
+gradient-based local optimisers (conjugate gradient), similarly outperformed
+purely stochastic methods \cite{doi:10.1002/qua.560440214}. Judson \textit{et
+al.} classify points on the model energy surface as being in one of three
+states: knotted and of high energy, unknotted and of low but not minimal energy,
+and unknotted with globally minimised energy \cite{doi:10.1002/qua.560440214}.
+They applied both GA and SA in order to make the transition from the first state
+to the second state before applying local optimisation to descend to the bottom
+of the basin.
 
-The Conformational Space Annealing Method is a more recent unbiased global
-optimisation algorithm that has produced very good results
-\cite{PhysRevLett.91.080201}. This technique is highly robust against random
-initial configurations thanks to its generality. As a result, it has therefore
-been able to produce all global minima for $n\leq183$ without suffering from
-exceptional cases like those experienced by the biased modelling algorithms.
-This algorithm combines concepts of Monte Carlo with Minimisation, Genetic
-Algorithm and simulated Annealing, and requires no specific knowledge or
-assumptions regarding the optimal structure. Local minimisation is performed on
-a set of configurations, and then similarly to the mating process of genetic
-algorithm, a subset of configurations is produced. A parameter $D_cut$ is
-introduced and acts as a temperature variable in the annealing process. This
-parameter represents a distance measure between configurations and is used to
-maintain diversity in the population \cite{PhysRevLett.91.080201}.
+Genetic operators for the two dimensional bonded Lennard-Jones problem have
+since been further developed with successful results \cite{PULLAN1998331},
+proving the GA approach to be a viable global optimisation technique. Pullan
+\cite{PULLAN1998331} proposes three genetic crossover methods and four mutation
+operators and applies these in a parallel genetic algorithm to find global
+minima for most values of $N, 2 \leq N \leq 55$, as well as a minimum for $N =
+61$ which was previously unseen in literature.
 
-Another unbiased global optimisation algorithm, known as dynamic lattice
-searching (DLS), works by repeatedly constructing and searching a dynamic
-lattice until a global minimum is produced \cite{doi:10.1002/jcc.20096}. This
-lattice is built by searching the possible location sites for added atoms, which
-are then iteratively relocated from lattice sites of higher energy to those of
-lower energy. This method is very efficient because of how much it reduces the
-search space, but remains unbiased and therefore does not exhibit the same
-limitations seen in the biased modelling based techniques.
+This two stage hybrid method has been successful, but Judson \textit{et al.}
+also suggest the possibility of adding a third optimisation stage which is
+intermediate between the global and local gradient methods
+\cite{doi:10.1002/qua.560440214}. It is thought that a non-gradient optimisation
+method such as the simplex method may provide a means of stepping over barriers
+into possibly deeper minima surrounding those found by the global method in the
+first phase.
 
-\section{Proposed Optimisation Algorithm} \label{proposed}
+More recently, the Conformational Space Annealing Method
+\cite{PhysRevLett.91.080201} has been developed as a global optimisation
+algorithm that has produced very good results when applied to the problem of an
+unbound cluster of atoms with Lennard-Jones relationships. This algorithm
+combines concepts of Monte Carlo with minimisation, genetic algorithms, and
+simulated annealing. While this technique has not been applied to the bonded
+variant of the Lennard-Jones conformation problem as discussed here, it
+demonstrates that a hybrid approach combining SA and evolutionary population
+based algorithms may be a viable technique in a similar problem domain.
+
+\section{Proposed Algorithm} \label{proposed}
+
+As discussed in section~\ref{litreview}, global meta-heuristics such as the
+genetic algorithm, have been used successfully as a means of navigating the
+global search space and avoiding local minima that are caused by knots in the
+configurations \cite{PULLAN1998331, doi:10.1002/qua.560440214}. These methods
+use the global optimiser to find unknotted potential candidate configurations
+and then refine them using gradient-based methods. Judson \textit{et al.} have
+suggested a three stage system in which a non-gradient optimiser such as the
+simplex method may be used as an intermediate stage between the global and
+local-gradient optimisation in order to traverse barriers into potentially
+deeper basins \cite{doi:10.1002/qua.560440214}.
+
+As simulated annealing has been found to be successful when combined with
+evolutionary algorithms in similar problem domains \cite{PhysRevLett.91.080201},
+this paper will explore its use as an intermediate non-gradient method such that
+the proposed algorithm has three levels: (1) global meta-heuristic, (2),
+intermediate non-gradient optimisation and (3) local gradient-based
+optimisation.
 

### 2018-08-16 documentation/paper.tex
Refactored angle->cartesian conversion into its own function since it'll be reused by the gradient.

@@
 
-\title{Lennard-Jones Energy Minimisation}
+\title{3 Stage Hybrid Optimisation Algorithm for the Bonded Lennard-Jones Molecular Conformation Problem}
 \author{Bennett Gale}
-\date{July 2018}
+\date{August 2018}
 \lstset{language=C++, basicstyle=\ttfamily}
@@
 The model molecular energy minimisation problem involves the geometric
-configuration of a cluster of atoms in two dimensions, each connected by rigid
+configuration of a chain of atoms in two dimensions, each connected by rigid
 bonds of unit length, such that the total potential energy yielded by the sum of
@@
 prevent premature convergence. Over many generations the population converges
-towards a global minimum as each configuration communicates
-information about the global search space through the crossover operation. This
-method has been successful in producing results that were previously unseen in
-literature for certain input sizes.
+towards a global minimum as each configuration communicates information about
+the global search space through the crossover operation. This method has been
+successful in producing results that were previously unseen in literature for
+certain input sizes \cite{PULLAN1998331}.
 
 Judson \textit{et al.} compare the performance of GAs with simulated annealing,
-a stochastic global optimisation technique in which Monte Carlo steps are taken
-and new states are accepted if they are fitter, or with a probability based on a
-function of a temperature value \cite{doi:10.1002/qua.560440214}. When applied
-to the molecular conformation problem, they found that the GA method and the SA
-method, both applied as global meta-heuristics in conjunction with
-gradient-based local optimisers (conjugate gradient), similarly outperformed
-purely stochastic methods \cite{doi:10.1002/qua.560440214}. Judson \textit{et
-al.} classify points on the model energy surface as being in one of three
-states: knotted and of high energy, unknotted and of low but not minimal energy,
-and unknotted with globally minimised energy \cite{doi:10.1002/qua.560440214}.
-They applied both GA and SA in order to make the transition from the first state
-to the second state before applying local optimisation to descend to the bottom
-of the basin.
+a probabilistic global optimisation technique in which Monte Carlo steps are
+taken and new states are accepted if they are fitter, or with a probability
+based on a function of a temperature value \cite{doi:10.1002/qua.560440214}.
+When applied to the molecular conformation problem, they found that the GA
+method and the SA method, both applied as global meta-heuristics in conjunction
+with gradient-based local optimisers (conjugate gradient), similarly
+outperformed purely stochastic methods \cite{doi:10.1002/qua.560440214}. Judson
+\textit{et al.} classify points on the model energy surface as being in one of
+three states: knotted and of high energy, unknotted and of low but not minimal
+energy, and unknotted with globally minimised energy
+\cite{doi:10.1002/qua.560440214}. They applied both GA and SA in order to make
+the transition from the first state to the second state before applying local
+optimisation to descend to the bottom of the basin.
 
@@
 this paper will explore its use as an intermediate non-gradient method such that
-the proposed algorithm has three levels: (1) global meta-heuristic, (2),
-intermediate non-gradient optimisation and (3) local gradient-based
-optimisation.
+the proposed algorithm has three levels: global meta-heuristic (GA),
+intermediate non-gradient optimisation (SA) and local gradient-based
+optimisation (BFGS). The performance of this three stage optimisation method
+will be compared with a purely evolutionary algorithm without the intermediate
+step.
+
+\subsection{Genetic Encoding}
+
+Encoding each  molecule as a chain of two-dimensional cartesian coordinates
+would require constraints to be enforced so that the relative orientations and
+unit length bonds are maintained. In order to avoid this, therefore, molecules
+will be encoded as sequences of $N-2$ angles, where $N$ is the number of atoms
+in the molecule and each angle, $\alpha_i$ represents the rotation of the
+segment following atom $a_i$. This encoding scheme means that cartesian
+coordinates must be calculated at each Lennard-Jones potential computation as it
+requires euclidean distances between pairs of atoms to be computed, but results
+in much simpler crossover and mutation operations.
+
+\subsection{Selection \& Crossover Method}
+
+An elitist methodology will be utilised throughout the GA optimisation, meaning
+that the single fittest configuration will be passed on unmodified to subsequent
+generations in order to preserve beneficial traits and provide a mechanism for
+remembering information about the best configuration that has been found so far.
+
+Pairs of parents to be crossed over will then be selected using the Roulette
+method, where the probability for a potential parent to be selected is based on
+a function of its fitness value. The `elite' configurations that were inherited
+from the previous generation will also be candidates for selection using this
+method. Selected parents will undergo one of the following crossover operators
+with equal probabilities of occurring:
+
+\subsubsection{Single-Point Crossover}
+
+A single atom $a_i$ will be randomly selected. Atoms to the left of this point
+in one parent will form the left segment of the child, and the right segment of
+the child will be similarly formed from the atoms to the right of the second
+parent. A second child may be generated by repeating with the opposite segments
+(the right segment from parent one and the left segment from parent two).
+Finally, the angle $\alpha_i$ at the crossover point will be iterated through
+discrete steps until an angle is found which results in the lowest energy, as
+described by Pullan \cite{PULLAN1998331}.
+
+\subsubsection{Two-Point Crossover}
+
+Two crossover points will be selected at random and the molecules will be
+combined similarly to the single-point crossover, with each child containing two
+segments from one parent and one segment from the other. As with the single
+point crossover, the angles of both crossover points will be optimised.
+
+\subsection{Mutation and Intermediate Optimisation}
+
+The mutation mechanism is where the intermediate non-gradient optimisation
+algorithm will be incorporated. In traditional genetic algorithms, mutation has
+a small chance of occurring and is random, resulting in some states which are
+less fit in the hopes that this will later lead to a deeper basin in the search
+space. By incorporating simulated annealing at the mutation stage it is hoped
+that the configuration will be dragged in a downward direction while also
+providing enough stochastic properties to allow boundaries between minima to be
+traversed, which is the goal of any mutation operator.
+
+Mutation will have a low chance of occurring so as not to corrupt fit
+configurations too frequently and lose information provided by the global
+search. Each generation, child nodes that are produced during the crossover
+stage will have a small chance of becoming initial states in the simulated
+annealing search.
+
+\subsubsection{Annealing Schedule}
+
+During annealing random angles are selected and modified by random amounts. If
+the change produces an improved state, it is accepted. If the change produces a
+worsened state, it is accepted with a probability of $1 - e^{\delta E / kT}$
+such that the current `temperature' as well as how bad the state is determines
+whether it is accepted.
+
+The goal at this intermediate stage is not to find the global minimum, but
+merely to provide a more localised search for deeper basins in the vicinity of
+the current configuration; it is hoped that if the GA search has found a local
+minimum in close proximity to the global minimum, the SA stage will take large
+enough steps in order to climb over the boundary into the neighbouring basin
+where it can be finally optimised by the gradient search in the next stage.
+
+Therefore, an annealing schedule that prioritises speed over optimality will be
+chosen.
+
+\subsection{Local Optimisation}
+
+The final stage of the algorithm, after a candidate configuration has been found
+by the GA and its immediate vicinity has been further explored through mutation,
+is to refine the configuration using gradient-based local search. This stage
+will use the gradient of the objective function in order to guarantee that the
+bottom of the current basin will be reached. The quasi-Newton method
+Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage resulting
+in a fast descent towards the current local minimum at each stage.
+
+The coordinates of this locally optimised configuration will also be passed back
+to the GA. This is known as a `Lamarkian' process and has been found to be
+beneficial when used with this problem \cite{doi:10.1002/qua.560440214}.
+
+%\subsection{Implementation}
+
+% \begin{algorithm} \caption{Proposed High-Level Algorithm} \label{algorithm}
+%   \centering \begin{algorithmic} \Procedure{optimise}{} % \State $nn \gets new
+%   Network$ %   \For{$i \gets 0; i <= number\_of\_epochs; i++$} %
+%   \For{$\textit{each data partition of minibatch size}$} %           \State
+%   $\textit{Do forward pass}$ %           \State $\textit{Backpropagate the
+%   resulting error}$ %           \State $\textit{Update the weights
+%   proportionally to their contribution to the error}$ %       \EndFor %
+%   \State $\textit{Shuffle and repartition the data}$ %   \EndFor \State $P
+%   \gets \textit{Initial Population Size}$ \State $population \gets initialise$
+%   \EndProcedure \end{algorithmic}
+
+% \end{algorithm}
 

### 2018-08-16 main.cpp

@@ -43,20 +43,26 @@ void print_csv(Molecule& s)
 	std::cout << std::endl;
 }
 
+
+void angle_to_cart(const Molecule& s, std::vector<Cartesian> &cart)
+{
+	cart[0] = {0, 0};
+	double phi = 0;
+	for (int i = 0; i < s.bond_angles.size(); i++) {
+		phi += s.bond_angles[i];
+		double x = cart[i].x + cos(phi);
+		double y = cart[i].y + sin(phi);
+		cart[i + 1] = {x, y};
+	}
+}
+
 /**
  * Calculate the Lennard-Jones potential energy of the given system s
  */
 double lj_potential(Molecule& s)
 {
-    std::vector<Cartesian> cart(s.length);
-    cart[0] = {0, 0};
-    double phi = 0;
-    for (int i = 0; i < s.bond_angles.size(); i++) {
-        phi += s.bond_angles[i];
-        double x = cart[i].x + cos(phi);
-        double y = cart[i].y + sin(phi);
-	    cart[i + 1] = {x, y};
-    }
+	std::vector<Cartesian> cart(s.length);
+	angle_to_cart(s, cart);
     double v = 0;
     for (int i = 0; i < s.length - 1; i++) {
         for (int j = i + 1; j < s.length; j++) {

### 2018-08-28 main.cpp
Added flatten mutation

@@ -9,6 +9,7 @@
 
 #define MAX_LBFGS_ITER 99
 #define MISC_BUF 100
+#define MUTATION_RATE 0.5f
 
 //#define DEBUG
 
@@ -16,9 +17,9 @@ struct Cartesian {
 	double x = 0, y = 0;
 };
 
-struct Fitness_pair {
-	double *mol;
-	double fitness;
+struct Candidate {
+	double *m;
+	double f;
 };
 
 struct Bfgs_molecule {
@@ -35,7 +36,6 @@ struct Bfgs_molecule {
 typedef void (*crossover_ptr)(
 	const double*,
 	const double*,
-	double*,
 	double*,
 	int,
 	double*,
@@ -188,8 +188,7 @@ int lbfgs_progress(
 void single_crossover(
 		const double *p1,
 		const double *p2,
-		double *c1,
-		double *c2,
+		double *c,
 		int n,
 		double *grads,
 		Cartesian *cart,
@@ -198,18 +197,38 @@ void single_crossover(
 )
 {
 	int cp = rand()%n;
-	std::copy(p1, p1+cp, c1);
-	std::copy(p2+cp, p2+n, c1+cp);
-	std::copy(p2, p2+cp, c2);
-	std::copy(p1+cp, p1+n, c2+cp);
+	std::copy(p1, p1+cp, c);
+	std::copy(p2+cp, p2+n, c+cp);
 
 	// Optimise at the crossover point
-	Bfgs_molecule bm = {c1, grads, cart, nullptr, n, n+1, 0, n};
+	int indices[1] = {cp};
+	Bfgs_molecule bm = {c, grads, cart, indices, n, n+1, 1, n};
 	lbfgsfloatval_t res;
-	lbfgs(n, c1, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
-	bm.molecule = c2;
-	lbfgs(n, c2, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
+	lbfgs(1, c+cp, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
 }
+
+//void single_mutation(double *m, int n, Cartesian *cart, int n_cart)
+//{
+//	if (n != n_cart - 1) {
+//		std::cerr << "Sizes don't match" << std::endl;
+//		exit(-1);
+//	}
+//	std::random_device rd;
[…]

### 2018-08-28 main.cpp
Added new mutation operators

@@ -207,30 +207,30 @@ void single_crossover(
 	lbfgs(1, c+cp, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
 }
 
-//void single_mutation(double *m, int n, Cartesian *cart, int n_cart)
-//{
-//	if (n != n_cart - 1) {
-//		std::cerr << "Sizes don't match" << std::endl;
-//		exit(-1);
-//	}
-//	std::random_device rd;
-//	std::mt19937 gen(rd()); // New Mersenne-Twister generator seeded with rd
-//	std::uniform_real_distribution<double> uniform(-M_PI, M_PI);
-//	angle_to_cart(m, cart, n, n_cart);
-//	double initial_e = lj_potential(cart, n);
-//	double e;
-//	do {
-//		int mp = rand() % n;
-//		double prev = m[mp];
-//		m[mp] = uniform(gen);
-//		angle_to_cart(m, cart, n, n_cart);
-//		e = lj_potential(cart, n_cart);
-//		if (e > initial_e) // Revert
-//			m[mp] = prev;
-//	} while (e > initial_e);
-//}
+void guided_single_mutation(double* m, int n, Cartesian* cart, int n_cart)
+{
+	if (n != n_cart - 1) {
+		std::cerr << "Sizes don't match" << std::endl;
+		exit(-1);
+	}
+	std::random_device rd;
+	std::mt19937 gen(rd()); // New Mersenne-Twister generator seeded with rd
+	std::uniform_real_distribution<double> uniform(-M_PI, M_PI);
+	angle_to_cart(m, cart, n, n_cart);
+	double initial_e = lj_potential(cart, n);
+	double e;
+	do {
+		int mp = rand() % n;
+		double prev = m[mp];
+		m[mp] = uniform(gen);
+		angle_to_cart(m, cart, n, n_cart);
+		e = lj_potential(cart, n_cart);
+		if (e > initial_e) // Revert
+			m[mp] = prev;
+	} while (e > initial_e);
+}
 
-void single_mutation(double *m, int n, Cartesian *cart, int n_cart)
+void stretch_mutation(double* m, int n, Cartesian* cart, int n_cart)
 {
 	if (n != n_cart - 1) {
 		std::cerr << "Sizes don't match" << std::endl;
@@ -251,7 +251,31 @@ void single_mutation(double *m, int n, Cartesian *cart, int n_cart)
 	}
 }
 
[…]

### 2018-08-30 main.cpp
Added double-point crossover

@@ -8,10 +8,13 @@
 #include <cstring>
 
 #define MAX_LBFGS_ITER 99
+#define MUTATION_CAP 10
 #define MISC_BUF 100
 #define MUTATION_RATE 0.5f
 
 //#define DEBUG
+
+const double optimal[] = {0, 0, -1.0, -3.0, -5.1, -7.2, -9.3, -12.5, -14.7, -16.9, -20.1, -22.3, -25.5, -27.8, -31.0,-33.2, -36.5, -38.7, -42.0, -45.3,-47.5, -50.8, -53.0, -56.3, -59.6,-61.8, -65.1, -68.4, -70.7, -74.0,-77.2, -79.5, -82.8, -86.1, -88.3,-91.7, -95.0, -98.3, -100.5, -103.8,-107.1, -109.4, -112.7, -116.0, -119.3,-121.6, -124.9, -128.6, -131.5, -133.8,-137.1, -140.5, -143.7, -146.0, -149.3,-152.7};
 
 struct Cartesian {
 	double x = 0, y = 0;
@@ -207,6 +210,33 @@ void single_crossover(
 	lbfgs(1, c+cp, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
 }
 
+void double_crossover(
+		const double *p1,
+		const double *p2,
+		double *c,
+		int n,
+		double *grads,
+		Cartesian *cart,
+		int n_grad,
+		int n_cart
+)
+{
+	int cpa = rand() % n;
+	int cpb = rand() % n;
+	if (cpa > cpb)
+		std::swap(cpa, cpb);
+	std::copy(p1, p1+cpa, c);
+	std::copy(p2+cpa, p2+cpb, c+cpa);
+	std::copy(p1+cpb, p1+n, c+cpb);
+
+	// Optimise at the crossover points
+	int indices[2] = {cpa, cpb};
+	double x[2] = {c[cpa], c[cpb]};
+	Bfgs_molecule bm = {c, grads, cart, indices, n, n+1, 2, n};
+	lbfgsfloatval_t res;
+	lbfgs(2, x, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
+}
+
 void guided_single_mutation(double* m, int n, Cartesian* cart, int n_cart)
 {
 	if (n != n_cart - 1) {
@@ -219,15 +249,16 @@ void guided_single_mutation(double* m, int n, Cartesian* cart, int n_cart)
[…]

### 2018-08-30 run.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+cmake-build-release/bonded_molecular_conformation | tee documentation/results.csv | ./plot.py

### 2018-08-30 main.cpp
added SA

@@ -36,7 +36,7 @@ struct Bfgs_molecule {
 	int n_gradient;
 };
 
-typedef void (*crossover_ptr)(
+typedef void (*Crossover_ptr)(
 	const double*,
 	const double*,
 	double*,
@@ -47,7 +47,7 @@ typedef void (*crossover_ptr)(
 	int
 );
 
-typedef void (*mutation_ptr)(double*, int, Cartesian*, int);
+typedef void (*Mutation_ptr)(double*, int, Cartesian*, int);
 
 // --------------------- Function Prototypes -------------------------- //
 double lj_potential(Cartesian *cart, int n);
@@ -210,34 +210,34 @@ void single_crossover(
 	lbfgs(1, c+cp, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
 }
 
-void double_crossover(
-		const double *p1,
-		const double *p2,
-		double *c,
-		int n,
-		double *grads,
-		Cartesian *cart,
-		int n_grad,
-		int n_cart
-)
-{
-	int cpa = rand() % n;
-	int cpb = rand() % n;
-	if (cpa > cpb)
-		std::swap(cpa, cpb);
-	std::copy(p1, p1+cpa, c);
-	std::copy(p2+cpa, p2+cpb, c+cpa);
-	std::copy(p1+cpb, p1+n, c+cpb);
+//void double_crossover(
+//		const double *p1,
+//		const double *p2,
+//		double *c,
+//		int n,
+//		double *grads,
+//		Cartesian *cart,
+//		int n_grad,
+//		int n_cart
+//)
+//{
+//	int cpa = rand() % n;
+//	int cpb = rand() % n;
+//	if (cpa > cpb)
+//		std::swap(cpa, cpb);
+//	std::copy(p1, p1+cpa, c);
+//	std::copy(p2+cpa, p2+cpb, c+cpa);
+//	std::copy(p1+cpb, p1+n, c+cpb);
+//
+//	// Optimise at the crossover points
+//	int indices[2] = {cpa, cpb};
+//	double x[2] = {c[cpa], c[cpb]};
+//	Bfgs_molecule bm = {c, grads, cart, indices, n, n+1, 2, n};
+//	lbfgsfloatval_t res;
+//	lbfgs(2, x, &res, lbfgs_evaluate, lbfgs_progress, &bm, nullptr);
+//}
 
-	// Optimise at the crossover points
-	int indices[2] = {cpa, cpb};
-	double x[2] = {c[cpa], c[cpb]};
[…]

### 2018-08-30 documentation/paper.tex
Implemented command line args

@@
 
-\title{3 Stage Hybrid Optimisation Algorithm for the Bonded Lennard-Jones Molecular Conformation Problem}
+\title{Evolutionary Algorithm for the Bonded Molecular Conformation Problem}
 \author{Bennett Gale}
@@
 Genetic algorithms have been applied \cite{doi:10.1002/qua.560440214,
-PULLAN1998331} successfully in order to extend the range of input size $N$ that
+PULLAN1998331, } successfully in order to extend the range of input size $N$ that
 is reachable through structural optimisation. This population based approach

### 2018-08-30 main.cpp

@@ -6,6 +6,7 @@
 #include <ctime>
 #include <numeric>
 #include <cstring>
+#include <cinttypes>
 
 #define MAX_LBFGS_ITER 99
 #define MUTATION_CAP 10
@@ -427,20 +428,23 @@ void ga(
 	delete [] grads;
 }
 
-int main()
+int main(int argc, char **argv)
 {
-	const int n_angles = 29;
-	const int n_cart = n_angles + 1;
-	const int n_gens = 100;
+	if (argc < 3) {
+		std::cout << "Usage: bonded_molecular_conformation <n> <number of generations>"
+			      << std::endl;
+		exit(-1);
+	}
+	const int n_cart = (int)std::strtol(argv[1], nullptr, 0);
+	const int n_angles = n_cart - 1;
+	const int n_gens = (int)std::strtol(argv[2], nullptr, 0);
 	srand((unsigned int) time(nullptr));
 	Crossover_ptr cross[1] = {single_crossover}; //double_crossover};
-	Mutation_ptr mut[3] = {sa_mutation, guided_mutation, stretch_mutation};
+	Mutation_ptr mut[3] = {naive_single_mutation, guided_mutation, stretch_mutation};
 	double sol[n_angles];
-
-	ga(cross, mut, sol, 12, 1, 3, n_angles, n_gens);
-	Cartesian cart[n_cart];
-	angle_to_cart(sol, cart, n_angles, n_cart);
-	double e = lj_potential(cart, n_cart);
+	for (int i = 0; i < 3; i++) {
+		ga(cross, mut, sol, 12, 1, 3, n_angles, n_gens);
+	}
 
 	return 0;
 }

### 2018-08-30 plot.py

@@ -1,6 +1,6 @@
 #!/usr/bin/env python3
 
-from sys import stdin
+from sys import stdin, argv
 import matplotlib.pyplot as plt
 from math import cos, sin
 
@@ -23,15 +23,19 @@ def plot(angle_chain, fig_name):
         y.append(y[-1] + sin(cumulative))
     plt.plot(x, y, 'o-')
     plt.title(chart_title)
-    plt.savefig("./documentation/figure " + fig_name + ".png")
+    plt.savefig(fig_name)
     plt.close()
-    print(x)
-    print(y)
+    # print(x)
+    # print(y)
 
 
 if __name__ == "__main__":
     count = 0
     for line in stdin:
         print(line)
-        plot(line, str(count))
+        if argv[1] != '':
+            figname = argv[1] + '/' + "Figure " + str(count)
+        else:
+            figname = "Figure " + str(count)
+        plot(line, figname)
         count += 1

### 2018-08-30 run.sh

@@ -1,3 +1,11 @@
 #!/usr/bin/env bash
 
-cmake-build-release/bonded_molecular_conformation | tee documentation/results.csv | ./plot.py
+NUM_GENS=100
+
+for i in {3..61};
+do
+    echo "Beginning N = $i"
+    filepath=documentation/without_sa/n$i
+    mkdir -p $filepath
+    cmake-build-release/bonded_molecular_conformation $i $NUM_GENS | tee $filepath/results.csv | ./plot.py $filepath
+done

### 2018-08-31 documentation/paper.tex
Tweaked SA parameters

@@
 Genetic algorithms have been applied \cite{doi:10.1002/qua.560440214,
-PULLAN1998331, } successfully in order to extend the range of input size $N$ that
-is reachable through structural optimisation. This population based approach
-emulates the Darwinian process of natural selection, and works by performing
-crossover operations on pairs of candidate configurations with selection
-probabilities based on their fitness, producing new child nodes which contain
-genetic material from both parents. With every crossover operation there is a
-small chance of subsequent random mutation of the child configurations, the
+PULLAN1998331, } successfully in order to extend the range of input size $N$
+that is reachable through structural optimisation. This population based
+approach emulates the Darwinian process of natural selection, and works by
+performing crossover operations on pairs of candidate configurations with
+selection probabilities based on their fitness, producing new child nodes which
+contain genetic material from both parents. With every crossover operation there
+is a small chance of subsequent random mutation of the child configurations, the
 purpose of which is to promote genetic diversity within the population and
@@
 
-Encoding each  molecule as a chain of two-dimensional cartesian coordinates
+Encoding each  molecule as a chain of two-dimensional Cartesian coordinates
 would require constraints to be enforced so that the relative orientations and
@@
 in the molecule and each angle, $\alpha_i$ represents the rotation of the
-segment following atom $a_i$. This encoding scheme means that cartesian
+segment following atom $a_i$. This encoding scheme means that Cartesian
 coordinates must be calculated at each Lennard-Jones potential computation as it
@@
 
-\subsection{Selection \& Crossover Method}
+\subsection{Selection \& Crossover Methods}
 
-An elitist methodology will be utilised throughout the GA optimisation, meaning
-that the single fittest configuration will be passed on unmodified to subsequent
-generations in order to preserve beneficial traits and provide a mechanism for
-remembering information about the best configuration that has been found so far.
+% An elitist methodology will be utilised throughout the GA optimisation,
+% meaning that the single fittest configuration will be passed on unmodified to
+% subsequent generations in order to preserve beneficial traits and provide a
+% mechanism for remembering information about the best configuration that has
+% been found so far.
 
-Pairs of parents to be crossed over will then be selected using the Roulette
-method, where the probability for a potential parent to be selected is based on
-a function of its fitness value. The `elite' configurations that were inherited
-from the previous generation will also be candidates for selection using this
-method. Selected parents will undergo one of the following crossover operators
-with equal probabilities of occurring:
+% Pairs of parents to be crossed over will then be selected using the Roulette
+% method, where the probability for a potential parent to be selected is based
+% on a function of its fitness value. The `elite' configurations that were
+% inherited from the previous generation will also be candidates for selection
+% using this method. Selected parents will undergo one of the following
+% crossover operators with equal probabilities of occurring:
 
-\subsubsection{Single-Point Crossover}
+The population is represented as a large pool of size $n_{pop} \times 2$ where
+$n_{pop}$ is the user-provided population size. This pool contains both the
+current population and the generated children in a single array such that it
+can be easily sorted based on fitness.
 
-A single atom $a_i$ will be randomly selected. Atoms to the left of this point
-in one parent will form the left segment of the child, and the right segment of
-the child will be similarly formed from the atoms to the right of the second
-parent. A second child may be generated by repeating with the opposite segments
-(the right segment from parent one and the left segment from parent two).
-Finally, the angle $\alpha_i$ at the crossover point will be iterated through
-discrete steps until an angle is found which results in the lowest energy, as
-described by Pullan \cite{PULLAN1998331}.
+% \subsubsection{Single-Point Crossover}
 
-\subsubsection{Two-Point Crossover}
+% A single atom $a_i$ will be randomly selected. Atoms to the left of this point
+% in one parent will form the left segment of the child, and the right segment
+% of the child will be similarly formed from the atoms to the right of the
+% second parent. A second child may be generated by repeating with the opposite
+% segments (the right segment from parent one and the left segment from parent
+% two). Finally, the angle $\alpha_i$ at the crossover point will be iterated
+% through discrete steps until an angle is found which results in the lowest
+% energy, as described by Pullan \cite{PULLAN1998331}.
 
-Two crossover points will be selected at random and the molecules will be
-combined similarly to the single-point crossover, with each child containing two
-segments from one parent and one segment from the other. As with the single
-point crossover, the angles of both crossover points will be optimised.
+\subsubsection{Crossover}
 
-\subsection{Mutation and Intermediate Optimisation}
+A single-point crossover operator is applied as follows: A single angle index
+$a_i$ is selected at random. Angles to the left of this point in the first
+parent form the left segment of the child, and the right segment of the child is
+similarly formed from the atoms to the right of the second parent.
+
+\paragraph{Partial L-BFGS}
+
+Finally, $\alpha_i$ is locally optimised by applying BFGS with this single angle
+as its only optimisation parameter, with the rest of the angles fixed. Energy
+and gradients for use in the BFGS algorithm are computed using the entire
+molecule, while only the subset is optimised, leaving the rest of the angles
+unchanged. This optimisation of the single crossover point effectively locally
+optimises the relative orientation of the two connected segments.
+%TODO: Should I leave this paragraph in?
+Crossover operators with multiple crossover-points were not found to be
+beneficial during experimentation and were even detrimental in some cases. As
+such, they have been excluded from this study. The focus is primarily on
+evaluating the use of annealing as a mutation operator and therefore the
+crossover operation has been kept simple.
+
+\subsubsection{Selection}
+
+Pairs of parent configurations to be crossed over are selected by choosing
+randomly from the fittest half of the population. This is repeated $n_{pop}$
+times so that the number of children created is equal to the population size and
+forms the next generation. Selecting from the fitter half of the population
+ensures that information about the global search space is retained through
+multiple generations by providing bias towards states that are closer to the
+global minimum.
+
+At each generation the entire pool is sorted so that the $n_{pop}$ fittest
+states, including children and fitter parents from the current generation move
+to the top half of the pool. This forms the new generation's population, and
+implicitly provides a form of elitism by allowing states from the previous
+generation to remain in the population if they are fit enough. After sorting,
+the second half of the pool contains discarded states which can be later
+overwritten with new children during subsequent crossover stages.
+
+\subsection{Mutation}
 
@@
 
-Mutation will have a low chance of occurring so as not to corrupt fit
+\subsubsection{Mutation Rate}
+
+Mutation generally has a low chance of occurring so as not to corrupt fit
 configurations too frequently and lose information provided by the global
-search. Each generation, child nodes that are produced during the crossover
-stage will have a small chance of becoming initial states in the simulated
-annealing search.
+search. It was found, however, that with a low mutation rate the search tended
+to converge prematurely due to an overly homogeneous population. In order to
+overcome this, the chance of mutation for each configuration is based on its
+similarity to its direct neighbour. That is, each state is mutated if its
+fitness is within one unit of that of the previous state.
 
-\subsubsection{Annealing Schedule}
+Additionally, if a state's energy is positive it is guaranteed to be mutated.
+This has the effect of reducing the number of knotted configurations in the
+population, as the mutation operators tend to guide them towards a negative
+energy state.
 
-During annealing random angles are selected and modified by random amounts. If
-the change produces an improved state, it is accepted. If the change produces a
-worsened state, it is accepted with a probability of $1 - e^{\delta E / kT}$
-such that the current `temperature' as well as how bad the state is determines
-whether it is accepted.
+\subsubsection{Mutation Operators}
 
-The goal at this intermediate stage is not to find the global minimum, but
-merely to provide a more localised search for deeper basins in the vicinity of
-the current configuration; it is hoped that if the GA search has found a local
-minimum in close proximity to the global minimum, the SA stage will take large
-enough steps in order to climb over the boundary into the neighbouring basin
-where it can be finally optimised by the gradient search in the next stage.
+\paragraph{Unravelling}
 
-Therefore, an annealing schedule that prioritises speed over optimality will be
-chosen.
+This mutation operator is based on one described by Pullan \cite{PULLAN1998331}
+in which A random angle index $\alpha_i$ is selected in the molecule and the
+distance between this point and each endpoint is compared. The angles between
+the selected angle and the closest end forms a new segment of angles. For each
+angle in the selected segment, a very small (close to zero) random number is
+chosen and the angle is set to that value. This results in one end of the
+molecule being `unravelled' into a straight line, which effectively removes any
+knots or kinks within that segment.The segment is later `rewound' through
+subsequent local optimisation.
+
+\paragraph{Guided Stochastic Mutation}
+
+A random angle is selected and the current fitness and the current value of this
+	angle is stored.The selected angle is set to a random value between $-\pi$
+	and $\pi$ and the fitness is recomputed. If the new fitness is less than the
+	previous fitness, revert the angle to its previous state. This process is
+	then repeated until the maximum number of iterations is reached. This limit
+	is important because after many generations it becomes increasingly
+	difficult to find a random change in an angle that does not give lower
+	energy, which would otherwise result in an infinite loop.
+
+\paragraph{Naive Mutation}
+
+\paragraph{Simulated Annealing Mutation}
+
+A relatively brief annealing schedule is used due to the fact that during every
+iteration an energy computation is required. The goal at this intermediate
+search algorithm is not to find the global minimum, but merely to provide a more
+localised search for deeper basins in the vicinity of the current configuration;
+it is hoped that if the GA search has found a local minimum in close proximity
+to the global minimum, the SA stage will take large enough steps in order to
+climb over the boundary into the neighbouring basin where it can be finally
+optimised by the gradient search in the next stage. Therefore, an annealing
+schedule that prioritises speed over optimality will be chosen. The initial
+temperature $T$ will be set to 100 and decrement by one every iteration until it
+reaches 0.
+
+During annealing, random angles are selected and modified by random amounts. If
+the change produces an improved state, it is always accepted. If the change
+produces a worsened state, it is accepted with a probability of $1 - e^{-\Delta
+E / Tk}$ such that both the current `temperature' as well as the degree to which
+the change will worsen the configuration, determine whether the change is to be
+accepted. $\Delta E$ represents the change in energy that will occur if the new
+configuration is accepted, $T$ represents the current temperature, and $k$ is
+used to scale probability curve. A value of $k=0.03$ was chosen as this provides
+a reasonable probability curve for energy deltas of 1 and below.
 
@@
 
-The final stage of the algorithm, after a candidate configuration has been found
-by the GA and its immediate vicinity has been further explored through mutation,
-is to refine the configuration using gradient-based local search. This stage
-will use the gradient of the objective function in order to guarantee that the
-bottom of the current basin will be reached. The quasi-Newton method
-Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage resulting
-in a fast descent towards the current local minimum at each stage.
+% The final stage of the algorithm, after a candidate configuration has been
+% found by the GA and its immediate vicinity has been further explored through
+% mutation, is to refine the configuration using gradient-based local search.
+% This stage will use the gradient of the objective function in order to
+% guarantee that the bottom of the current basin will be reached. The
+% quasi-Newton method Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at
+% this stage resulting in a fast descent towards the current local minimum at
+% each stage.
 
-The coordinates of this locally optimised configuration will also be passed back
-to the GA. This is known as a `Lamarkian' process and has been found to be
-beneficial when used with this problem \cite{doi:10.1002/qua.560440214}.
+% The coordinates of this locally optimised configuration will also be passed
+% back to the GA. This is known as a `Lamarkian' process and has been found to
+% be beneficial when used with this problem \cite{doi:10.1002/qua.560440214}.
+
+\subsection{GA Procedure and Parameters}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Appendix}
 

### 2018-08-31 main.cpp

@@ -328,7 +328,7 @@ void sa_mutation(double *m, int n, Cartesian *cart, int n_cart)
 		angle_to_cart(m, cart, n, n_cart);
 		double e_prime = lj_potential(cart, n);
 		double r = uniform_percentage(gen);
-		if (e_prime < e || exp(-(e_prime - e)/T) > r) {
+		if (e_prime < e || exp(-(e_prime - e)/T*0.03) > r) {
 			e = e_prime;
 		} else {
 			m[mp] = prev; // revert
@@ -356,6 +356,7 @@ void ga(
 	auto cart_buf = new Cartesian[n_angles + 1];
 	double global_best = std::numeric_limits<double>::max();
 	int best_gen = 0;
+	double best_e = std::numeric_limits<double>::max();
 	Bfgs_molecule bm;
 	for (int i = 0; i < p; i++) {
 		pop[i].m = new double[n_angles];
@@ -408,6 +409,11 @@ void ga(
 			return a.f < b.f;
 		});
 
+		if (pop[0].f < best_e) {
+			best_e = pop[0].f;
+			best_gen = g;
+		}
+
 		if (fabs(optimal[n_angles+1] - pop[0].f) <= 0.01)
 			break;
 	}
@@ -415,9 +421,11 @@ void ga(
 	// Return the best solution
 	std::copy(pop[0].m, pop[0].m + n_angles, sol);
 
-	for (int i = 0; i < n_pop; i++) {
+	for (int i =0; i < n_pop; i++) {
 		print_csv(pop[i].m, pop[i].f, n_angles);
 	}
+
+	std::cout << best_gen << std::endl;
 
 	// cleanup
 	for (int i = 0; i < n_pop; i++)
@@ -440,11 +448,9 @@ int main(int argc, char **argv)
 	const int n_gens = (int)std::strtol(argv[2], nullptr, 0);
 	srand((unsigned int) time(nullptr));
 	Crossover_ptr cross[1] = {single_crossover}; //double_crossover};
-	Mutation_ptr mut[3] = {naive_single_mutation, guided_mutation, stretch_mutation};
+	Mutation_ptr mut[3] = {sa_mutation, guided_mutation, stretch_mutation};
 	double sol[n_angles];
-	for (int i = 0; i < 3; i++) {
-		ga(cross, mut, sol, 12, 1, 3, n_angles, n_gens);
-	}
+	ga(cross, mut, sol, 12, 1, 3, n_angles, n_gens);
 
 	return 0;
 }

### 2018-08-31 run.sh

@@ -2,10 +2,10 @@
 
 NUM_GENS=100
 
-for i in {3..61};
+for i in {51..61};
 do
     echo "Beginning N = $i"
-    filepath=documentation/without_sa/n$i
+    filepath=documentation/with_sa_gens_kvar/n$i
     mkdir -p $filepath
     cmake-build-release/bonded_molecular_conformation $i $NUM_GENS | tee $filepath/results.csv | ./plot.py $filepath
 done

### 2018-08-31 documentation/paper.tex
Added tables and figures

@@
 \maketitle
-\tableofcontents
-\pagebreak
+% \tableofcontents
+% \pagebreak
 
@@
 optimises the relative orientation of the two connected segments.
+
 %TODO: Should I leave this paragraph in?
@@
 
-\subsubsection{Mutation Rate}
+\subsubsection{Mutation Rate} \label{mutationrate}
 
@@
 
+The following operators mutate a given configuration in both biased and
+unbiased ways.
+
 \paragraph{Unravelling}
@@
 
-A random angle is selected and the current fitness and the current value of this
-	angle is stored.The selected angle is set to a random value between $-\pi$
-	and $\pi$ and the fitness is recomputed. If the new fitness is less than the
-	previous fitness, revert the angle to its previous state. This process is
-	then repeated until the maximum number of iterations is reached. This limit
-	is important because after many generations it becomes increasingly
-	difficult to find a random change in an angle that does not give lower
-	energy, which would otherwise result in an infinite loop.
+A random angle is selected and both the current fitness of the molecule and the
+current value of this angle is stored.The selected angle is set to a random
+value between $-\pi$ and $\pi$ and the fitness is recomputed. If the new fitness
+is worse than the previous fitness, the angle is reverted to its previous state.
+This process is then repeated until either the maximum number of iterations is
+reached, or an improved configuration is found. This limit is important because
+after many generations it becomes increasingly difficult to find a random change
+in an angle that does not give lower energy, which would otherwise result in an
+infinite loop.
 
@@
 
-\paragraph{Simulated Annealing Mutation}
+This is an unbiased mutation operator which simply selects a random angle
+$\alpha_i$ and sets it to a random new value between $-\pi$ and $\pi$. This
+operator has no regard for whether the mutation improves or worsens state, and
+primarily serves the purpose of reducing homogeneity and premature convergence
+within the population.
+
+\paragraph{Annealed Mutation}
 
@@
 
-\subsection{Local Optimisation}
+\subsection{GA Procedure and Parameters}
 
-% The final stage of the algorithm, after a candidate configuration has been
-% found by the GA and its immediate vicinity has been further explored through
-% mutation, is to refine the configuration using gradient-based local search.
-% This stage will use the gradient of the objective function in order to
-% guarantee that the bottom of the current basin will be reached. The
-% quasi-Newton method Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at
-% this stage resulting in a fast descent towards the current local minimum at
-% each stage.
+The following GA parameters are used:
+\begin{description}
+	\item[Population Size] 12
+	\item[Maximum Generations] 100
+	\item[Mutation Rate] Deterministic and based on conditions as described in
+	section~\ref{mutationrate}
+\end{description}
+The GA works by continually sorting a large pool (of size $n_{pop} \times 2$) of
+configurations in order to rank them in order of fitness and subsequently select
+states for crossover and mutation operators. The population is initialised with
+random molecules and sorted and then the procedure outlined in REF are carried
+out until either a known global optimum is found or the maximum number of
+generations is reached.
 
-% The coordinates of this locally optimised configuration will also be passed
-% back to the GA. This is known as a `Lamarkian' process and has been found to
-% be beneficial when used with this problem \cite{doi:10.1002/qua.560440214}.
+%%%% GA Pseudocode here
 
-\subsection{GA Procedure and Parameters}
+The final stage of each GA iteration, after candidate configurations have been
+found by the GA and its immediate vicinity has been further explored through
+mutation, is to refine the configuration using gradient-based local search. This
+stage uses the gradient of the objective function in order to guarantee that the
+bottom of the current basin will be reached. The quasi-Newton method
+Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage resulting
+in a fast descent towards the current local minimum at each stage. The
+coordinates of this locally optimised configuration are also be passed back to
+the GA so that improvements are retained in future iterations. This is known as
+a `Lamarkian'process and has previously been found to be beneficial when used
+with this problem \cite{doi:10.1002/qua.560440214}.
 
@@
 
-\section{Discussion}
+\begin{figure}
+	\includegraphics[width=\textwidth]{nvse.pdf}
+	\caption{Lennard-Jones Potential as N size increases}
+	\label{fig:nvse}
+\end{figure}
+
+\section{Conclusion}
 
 \section{Appendix}
+
+\begin{table}[]
+	\label {tab:nosa}
+	\centering
+	\begin{tabular}{cccc|cccc}
+	$N$  & Optimal & Found & Gen. & $N$  & Optimal & Found  & Gen. \\
+	\hline
+	3  & -3.0    & -3.0  & 0    & 30 & -77.3   & -77.2  & 89   \\
+	4  & -5.1    & -5.1  & 1    & 31 & -79.5   & -79.6  & 70   \\
+	5  & -7.2    & -7.2  & 1    & 32 & -82.9   & -82.9  & 88   \\
+	6  & -9.3    & -9.4  & 52   & 33 & -86.1   & -86.2  & 87   \\
+	7  & -12.5   & -12.5 & 67   & 34 & -88.4   & -88.5  & 99   \\
+	8  & -14.7   & -14.7 & 75   & 35 & -91.7   & -90.7  & 99   \\
+	9  & -16.9   & -16.9 & 0    & 36 & -95.0   & -94.1  & 89   \\
+	10 & -20.1   & -20.1 & 0    & 37 & -98.3   & -97.3  & 98   \\
+	11 & -22.3   & -22.3 & 99   & 38 & -100.6  & -100.6 & 88   \\
+	12 & -25.6   & -25.6 & 96   & 39 & -103.8  & -103.7 & 86   \\
+	13 & -27.8   & -27.8 & 21   & 40 & -107.2  & -106.0 & 79   \\
+	14 & -31.0   & -31.0 & 95   & 41 & -109.4  & -108.5 & 96   \\
+	15 & -33.2   & -33.3 & 78   & 42 & -112.8  & -112.7 & 98   \\
+	16 & -36.5   & -36.5 & 5    & 43 & -116.0  & -114.1 & 97   \\
+	17 & -38.7   & -38.7 & 79   & 44 & -119.4  & -117.3 & 93   \\
+	18 & -42.0   & -42.0 & 77   & 45 & -121.6  & -120.6 & 96   \\
+	19 & -45.3   & -45.3 & 87   & 46 & -125.0  & -122.9 & 87   \\
+	20 & -47.5   & -47.6 & 76   & 47 & -128.6  & -125.3 & 94   \\
+	21 & -50.8   & -50.8 & 89   & 48 & -131.6  & -129.5 & 95   \\
+	22 & -53.0   & -53.1 & 99   & 49 & -133.8  & -130.4 & 78   \\
+	23 & -56.3   & -56.4 & 74   & 50 & -137.2  & -134.0 & 93   \\
+	24 & -59.6   & -59.6 & 90   & 51 & -140.5  & -137.4 & 72   \\
+	25 & -61.9   & -61.9 & 91   & 52 & -143.7  & -140.7 & 95   \\
+	26 & -65.1   & -65.1 & 22   & 53 & -146.1  & -142.6 & 98   \\
+	27 & -68.4   & -67.4 & 97   & 54 & -149.3  & -147.0 & 99   \\
+	28 & -70.7   & -70.7 & 53   & 55 & -152.7  & -150.0 & 68   \\
+	29 & -74.0   & -74.0 & 77   &    &         &        &      \\
+
+	\hline 
+	\end{tabular}
+	\caption{Results with unravelling, guided stochastic, and stochastic
+	mutation
+	operators}
+ \end{table}
+
+ \begin{table}[]
+	\label {tab:yessa}
+	\centering
+	\begin{tabular}{cccc|cccc}
+	$N$  & Optimal & Found & Gen. & $N$  & Optimal & Found  & Gen. \\
+	\hline
+	3  & -3.0    & -3     &    0   & 30 & -77.3   & -76.3    &  89  \\
+	4  & -5.1    & -5.1   &    0   & 31 & -79.5   & -78.5    &  81  \\
+	5  & -7.2    & -7.2   &    23  & 32 & -82.9   & -82.9    &  95  \\
+	6  & -9.3    & -9.4   &    90  & 33 & -86.1   & -86.2    &  98  \\
+	7  & -12.5   & -12.5  &    90  & 34 & -88.4   & -86.2    &  93  \\
+	8  & -14.7   & -14.7  &    93  & 35 & -91.7   & -90.8    &  99  \\
+	9  & -16.9   & -16.9  &    2   & 36 & -95.0   & -94.0    &  95  \\
+	10 & -20.1   & -20.1  &    1   & 37 & -98.3   & -96.2    &  86  \\
+	11 & -22.3   & -22.3  &    53  & 38 & -100.6  & -98.5    &  90  \\
+	12 & -25.6   & -25.6  &    80  & 39 & -103.8  & -102.7   &  94  \\
+	13 & -27.8   & -27.8  &    24  & 40 & -107.2  & -104.2   &  61  \\
+	14 & -31.0   & -31.0  &    56  & 41 & -109.4  & -108.5   &  87  \\
+	15 & -33.2   & -33.23 &    73  & 42 & -112.8  & -111.7   &  95  \\
+	16 & -36.5   & -36.5  &    15  & 43 & -116.0  & -112.8   &  90  \\
+	17 & -38.7   & -38.8  &    72  & 44 & -119.4  & -115.2   &  92  \\
+	18 & -42.0   & -42.1  &    93  & 45 & -121.6  & -117.3   &  99  \\
+	19 & -45.3   & -45.3  &    94  & 46 & -125.0  & -123.0   &  96  \\
+	20 & -47.5   & -47.5  &    86  & 47 & -128.6  & -126.3   &  88  \\
+	21 & -50.8   & -50.8  &    22  & 48 & -131.6  & -127.2   &  77  \\
+	22 & -53.0   & -53.0  &    61  & 49 & -133.8  & -130.7   &  67  \\
+	23 & -56.3   & -56.4  &    92  & 50 & -137.2  & -134.0   &  83  \\
+	24 & -59.6   & -58.7  &    76  & 51 & -140.5  & -134.2   &  98  \\
+	25 & -61.9   & -61.9  &    70  & 52 & -143.7  & -138.3   &  98  \\
+	26 & -65.1   & -65.1  &    3   & 53 & -146.1  & -142.9   &  45  \\
+	27 & -68.4   & -67.4  &    79  & 54 & -149.3  & -146.3   &  85  \\
+	28 & -70.7   & -70.7  &    67  & 55 & -152.7  & -149.5   &  97  \\
+	29 & -74.0   & -73.0  &    90  &    &         &          &       \\
+
+	\hline 
+	\end{tabular}
+	\caption{Results with unravelling, guided stochastic, and annealed mutation
+	operators}
+ \end{table}
+
+ \begin{figure}
+	\includegraphics[width=\textwidth]{figs/combined1.png}
+	\caption{Best Solution Found for N=5 to N=30}
+	\label{fig:combined1}
+\end{figure}
+
+ \begin{figure}
+	\includegraphics[width=\textwidth]{figs/combined2.png}
+	\caption{Best Solution Found for N=35 to N=55}
+	\label{fig:combined2}
+\end{figure}
 

### 2018-08-31 main.cpp

@@ -380,7 +380,7 @@ void ga(
 			return a.f < b.f;
 		});
 
-		// Mutate all children with a small chance
+		// Mutate
 		for (int i = 1; i < p; i++) {
 			if (fabs(pop[i].f - pop[i-1].f) < 1.0 || pop[i].f > 0)
 				mut[rand() % n_mut](pop[i].m, n_angles, cart_buf, n_angles+1);

### 2018-08-31 documentation/paper.tex
Finished paper

@@
 \usepackage{listings}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
 \usepackage{amsmath}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
 
@@
 individual Lennard-Jones potentials is globally minimised. The total potential
-energy of the cluster is given by
+energy of the system is given by
 $$V(r)=\sum^{N-1}_{i=1}\sum^{N}_{j=i+1}\left(r^{-12}_{ij}-2r^{-6}_{ij}\right)$$
@@
 
-The energy minimisation problem is historically difficult because of the
+The bound energy minimisation problem is historically difficult because of the
 exponential relationship between the locally stable structures and the number of
@@
 100, there exist more than $10^{140}$ minima \cite{DAVEN1996195}. It is
-therefore necessary to employ sophisticated search algorithms in order to make
+therefore necessary to employ global search algorithms in order to make
 larger input sizes attainable.
 
-Section~\ref{litreview} will discuss some of the research that has already taken
+The primary difficulty encountered with the bonded variation of the problem (as
+opposed to the unbound `Lennard-Jones cluster' problem), is that when atomic
+bonds overlap, a large amount of energy is required to pull an atom through the
+knot \cite{doi:10.1002/qua.560440214}. This is particularly problematic when
+local optimisation algorithms are applied, as they will not move towards the
+higher energy states required for knots to be removed. Due to this limitation of
+local search methods, a global algorithm is required in order to facilitate the
+traversal between local minima in the global search space by effectively finding
+unknotted configurations.
+
+Section~\ref{litreview} will discuss relevant research that has already taken
 place in this area. Section~\ref{proposed} will detail the algorithm proposed by
@@
 the proposed algorithm has three levels: global meta-heuristic (GA),
-intermediate non-gradient optimisation (SA) and local gradient-based
-optimisation (BFGS). The performance of this three stage optimisation method
-will be compared with a purely evolutionary algorithm without the intermediate
-step.
+intermediate non-gradient optimisation (SA) applied here in the form of a
+genetic
+mutation function, and local gradient-based optimisation (BFGS). The performance
+of this three stage optimisation method will be compared with a purely
+evolutionary algorithm without the intermediate step.
 
@@
 The mutation mechanism is where the intermediate non-gradient optimisation
-algorithm will be incorporated. In traditional genetic algorithms, mutation has
+algorithm is incorporated. In traditional genetic algorithms, mutation has
 a small chance of occurring and is random, resulting in some states which are
 less fit in the hopes that this will later lead to a deeper basin in the search
-space. By incorporating simulated annealing at the mutation stage it is hoped
+space. By incorporating simulated annealing at the mutation stage it is expected
 that the configuration will be dragged in a downward direction while also
@@
 This mutation operator is based on one described by Pullan \cite{PULLAN1998331}
-in which A random angle index $\alpha_i$ is selected in the molecule and the
+in which a random angle index $\alpha_i$ is selected in the molecule and the
 distance between this point and each endpoint is compared. The angles between
@@
 after many generations it becomes increasingly difficult to find a random change
-in an angle that does not give lower energy, which would otherwise result in an
-infinite loop.
+in an angle that does not increase the energy, which without the iteration limit
+would result in an infinite loop.
 
-\paragraph{Naive Mutation}
+\paragraph{Naive Stochastic Mutation}
 
@@
 A relatively brief annealing schedule is used due to the fact that during every
-iteration an energy computation is required. The goal at this intermediate
+iteration an energy computation is required. The goal of this intermediate
 search algorithm is not to find the global minimum, but merely to provide a more
@@
 produces a worsened state, it is accepted with a probability of $1 - e^{-\Delta
-E / Tk}$ such that both the current `temperature' as well as the degree to which
+E / Tk}$ such that both the current `temperature', as well as the degree to which
 the change will worsen the configuration, determine whether the change is to be
@@
 \end{description}
+This study compares the performance of the GA using an annealed mutation
+operator with that using only a purely stochastic operator. Both tests also use
+the guided mutation and the `unravel' mutation as possible operators to be
+selected at random.
+
 The GA works by continually sorting a large pool (of size $n_{pop} \times 2$) of
@@
 states for crossover and mutation operators. The population is initialised with
-random molecules and sorted and then the procedure outlined in REF are carried
-out until either a known global optimum is found or the maximum number of
-generations is reached.
+random molecules and sorted and then the procedure outlined in
+algorithm~\ref{pseudo} are carried out until either a known global optimum is
+found or the maximum number of generations is reached.
 
-%%%% GA Pseudocode here
-
+\begin{algorithm}[H] \label{pseudo}
+	initialise population\;
+	\For{$\leftarrow 0$ \KwTo $max\_gen$} {
+		\For{$c\leftarrow n\_pop \KwTo n\_pop * 2$} {
+			$p1\leftarrow Population[rand \% pop\_size / 2]$\;
+			$p2\leftarrow Population[rand \% pop\_size / 2]$\;
+			$Population[c] \leftarrow Crossover(p1, p2)$\;
+		}
+		\ForEach{s in Population} {
+			$s.fitness \leftarrow calculateFitness()$\;
+		}
+		$quicksort(Population)$\;
+		\For{$i \leftarrow 1 \KwTo n\_pop * 2$} {
+			\If {$Population[i] - Population[i-1] < 1.0 \textit{OR} Population
+			[i].fitness > 0$} {
+				$randomMutation(Population[i])$\;
+			}
+		}
+		\ForEach{s in Population} {
+			$BFGS(s)$\;
+		}
+		$quicksort(Population)$
+	}
+\end{algorithm}
 The final stage of each GA iteration, after candidate configurations have been
-found by the GA and its immediate vicinity has been further explored through
-mutation, is to refine the configuration using gradient-based local search. This
-stage uses the gradient of the objective function in order to guarantee that the
-bottom of the current basin will be reached. The quasi-Newton method
-Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage resulting
-in a fast descent towards the current local minimum at each stage. The
+found by the GA and their immediate vicinities have been further explored
+through mutation, is to refine the configuration using gradient-based local
+search. This stage uses the gradient of the objective function in order to
+guarantee that the bottom of the current basin will be reached. The quasi-Newton
+method Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage
+resulting in a fast descent towards the current local minimum at each stage. The
 coordinates of this locally optimised configuration are also be passed back to
@@
 
-\section{Results}
+\section{Results \& Conclusion}
 
@@
 
-\section{Conclusion}
+The genetic algorithm was tested in two configurations for comparison:
+\begin{itemize}
+	\item With the `unravel' mutation operator, the guided mutation operator and
+	the purely stochastic operator
+	\item With the `unravel' and guided mutation operators but replacing the
+	stochastic operator with the annealed operator.
+\end{itemize}
+It was expected that the annealed mutation would yield more optimal results
+for higher $N$ values by providing a means of traversing barriers between local
+minima and steeper basins within their vicinity. However, as seen in figure~
+\ref{fig:nvse}, while optimal results were found equally well by both
+configurations for around $n \leq 30$, at larger $n$ values the configuration
+with the annealed mutation operator performed less consistently and found
+global minima less often than the version with the stochastic mutation operator
+and no SA.
+
+The SA mutation operator was successful in its goal of extending the search to
+configurations
+surrounding those found by the GA, which was apparent in the increased variation
+within populations compared to the relatively homogeneous populations of the
+stochastic version. However, this increased variation may
+have been detrimental to the overall performance of the algorithm. Given the
+relatively high rate of mutation and the nature of the simulated annealing
+method, multiple worsened states are accepted for each mutation stage as opposed
+to the stochastic version which can only accept one worsened state per
+mutation. In other words, the simulated annealing approach is too aggressive
+for use as a mutation operator because early in the annealing process it accepts
+potentially many worsened states, throwing away important information found by
+the GA.
+
+It is thought that perhaps a less aggressive acceptance function or applying SA
+at a different stage in the algorithm (not as a mutation operator) may negate
+these effects, but it is apparent that the genetic algorithm approach provides
+sufficient global exploration, making the SA stage redundant.
 
@@
 
-%\subsection{Implementation}
-
-% \begin{algorithm} \caption{Proposed High-Level Algorithm} \label{algorithm}
-%   \centering \begin{algorithmic} \Procedure{optimise}{} % \State $nn \gets new
-%   Network$ %   \For{$i \gets 0; i <= number\_of\_epochs; i++$} %
-%   \For{$\textit{each data partition of minibatch size}$} %           \State
-%   $\textit{Do forward pass}$ %           \State $\textit{Backpropagate the
-%   resulting error}$ %           \State $\textit{Update the weights
-%   proportionally to their contribution to the error}$ %       \EndFor %
-%   \State $\textit{Shuffle and repartition the data}$ %   \EndFor \State $P
-%   \gets \textit{Initial Population Size}$ \State $population \gets initialise$
-%   \EndProcedure \end{algorithmic}
-
-% \end{algorithm}
-
 \pagebreak

### 2018-08-31 main.cpp

@@ -7,6 +7,7 @@
 #include <numeric>
 #include <cstring>
 #include <cinttypes>
+#include <algorithm>
 
 #define MAX_LBFGS_ITER 99
 #define MUTATION_CAP 10

### 2018-08-31 main.cpp
Removed calls to pow()

@@ -105,8 +105,12 @@ double lj_potential(Cartesian *cart, int n)
 		for (int j = i+1; j<n; j++) {
 			double delta_x = cart[j].x-cart[i].x;
 			double delta_y = cart[j].y-cart[i].y;
-			v += 1/pow(pow(delta_x, 2)+pow(delta_y, 2), 6)
-					-2/pow(pow(delta_x, 2)+pow(delta_y, 2), 3);
+			double dx_sqrd = delta_x * delta_x;
+			double dy_sqrd = delta_y * delta_y;
+			double dxdy_sqrd = dx_sqrd + dy_sqrd;
+			double dxdy_sqrd_3 = dxdy_sqrd * dxdy_sqrd * dxdy_sqrd;
+			double dxdy_sqrd_6 = dxdy_sqrd_3 * dxdy_sqrd_3;
+			v += 1/dxdy_sqrd_6 -2/dxdy_sqrd_3;
 		}
 	}
 	return v;
@@ -126,9 +130,9 @@ void lj_gradient(const Cartesian *cart, double *gradients, int n_cart, int n_gra
 				double dx = cart[i].x-cart[j].x;
 				double dy = cart[i].y-cart[j].y;
 				double r_sqrd = pow(dx, 2)+pow(dy, 2);
-				double res = (1.0/pow(r_sqrd, 7)-1.0/pow(r_sqrd, 4))
-						*(dx*(cart[m].y-cart[j].y)
-								+dy*(cart[j].x-cart[m].x));
+				double r_sqrd_4 = r_sqrd * r_sqrd * r_sqrd * r_sqrd;
+				double r_sqrd_7 = r_sqrd_4 * r_sqrd * r_sqrd * r_sqrd;
+				double res = (1.0/r_sqrd_7-1.0/r_sqrd_4) * (dx*(cart[m].y-cart[j].y) +dy*(cart[j].x-cart[m].x));
 				dv += res;
 			}
 		}

### 2018-08-31 main.cpp
Removed forgotten pow() call

@@ -55,9 +55,6 @@ typedef void (*Mutation_ptr)(double*, int, Cartesian*, int);
 double lj_potential(Cartesian *cart, int n);
 // -----------------------------------------------------------------------
 
-/**
- * This function is completely random and does not consider knots
- */
 void randomise_mol(double *molecule, int n)
 {
 	std::random_device rd;
@@ -129,7 +126,7 @@ void lj_gradient(const Cartesian *cart, double *gradients, int n_cart, int n_gra
 			for (int j = m+1; j<n_cart; j++) {
 				double dx = cart[i].x-cart[j].x;
 				double dy = cart[i].y-cart[j].y;
-				double r_sqrd = pow(dx, 2)+pow(dy, 2);
+				double r_sqrd = (dx * dx) + (dy * dy);
 				double r_sqrd_4 = r_sqrd * r_sqrd * r_sqrd * r_sqrd;
 				double r_sqrd_7 = r_sqrd_4 * r_sqrd * r_sqrd * r_sqrd;
 				double res = (1.0/r_sqrd_7-1.0/r_sqrd_4) * (dx*(cart[m].y-cart[j].y) +dy*(cart[j].x-cart[m].x));
@@ -430,7 +427,7 @@ void ga(
 		print_csv(pop[i].m, pop[i].f, n_angles);
 	}
 
-	std::cout << best_gen << std::endl;
+	//std::cout << best_gen << std::endl;
 
 	// cleanup
 	for (int i = 0; i < n_pop; i++)
